The Emergency Homeless Shelter Availability Across Toronto is Insufficient For Overwhelming Demand

### 2024-09-27 paper/paper.qmd
minor writing edits

Moving onto FSA-based observations, it was found that there are several neighbourhoods, including some around Old Toronto and North York, that report the highest concentrations of emergency shelters and the highest funding capacities (@tbl-fsa-numshelters). Despite the high occupancy in these specific areas, as shown in [@fig-regionoccupancy], it is important to remember that these findings are purely correlational and not causational. From this data alone, it is impossible to conclude whether these areas initially had greater rates of homelessness, thereby leading the government to spend more funds building shelters here, or that it is the higher concentration and availability of shelters here that will attract more homeless population to migrate here. Despite our lack of understanding of the two variable's true relationship, the correlation alone is able to indicate that emergency services are able to effectively reach the homeless population when they are placed here. When the city council is making policies and investments into emergency resources, it will be imperative to balance out their prioritization of these major areas of high demand versus areas which presently have less access, but may have comparatively lower demand.

@@

## Limitations and Next Steps {#sec-discussion-limitations}
One major limitation of the current dataset is that it only covers data from January to September 2024 as of the writing of this paper. This means that any patterns present in the winter months of November and December are not observable to the current analysis. As well, due to the dataset being unaudited, its accuracy to the true situations at these shelter locations cannot be fully verified.

@@

sample_data <- data %>% top_n(-5, id)
knitr::kable(sample_data, 
             align = "c",
             col.names = c("id", "Month", "Date", "FSA", "Location ID", 
                           "Actual Capacity", "Funding Capacity", 
                           "Occupied Beds", "Unoccupied Beds", 
                           "Unavailable Beds", "Occupancy Rate"))
```

>>>>

Moving onto FSA-based observations, it was found that there are several neighbourhoods, including some around Old Toronto and North York, that report the highest numbers of emergency shelter locations and the highest funding capacities (@tbl-fsa-numshelters). At first glance, the high occupancy also observed in the areas with higher funding seems to suggest that these regions experience more severe homelessness. However, it is important to remember that these findings are purely correlational and not causational. From this data alone, it is impossible to conclude whether these areas initially had greater rates of homelessness, thereby leading the government to spend more funds building shelters here, or that it is the higher concentration and availability of shelters here that will attract more homeless population to migrate here. Despite our lack of understanding of the two variables' true relationship, the correlation alone is able to indicate that emergency services are able to effectively reach the homeless population when they are placed here. When the city council is making policies and investments into emergency resources, it will be imperative to balance out their prioritization of these major areas of high demand versus areas which presently have less access, but may have comparatively lower demand.

@@

## Limitations and Next Steps {#sec-discussion-limitations}
One major limitation of the current dataset is that it only covers data from January to September 26, 2024 as of the writing of this paper. This means that any patterns present in the winter months of November and December are not observable to the current analysis. As well, due to the dataset being unaudited, its accuracy to the true situations at these shelter locations cannot be fully verified.

@@

sample_data <- data %>% top_n(-5, id) %>% 
  select(month, date, fsa, location_id, actual_capacity, funding_capacity, 
         occupied_beds, unoccupied_beds, unavailable_beds, occupancy_rate)
knitr::kable(sample_data, 
             align = "c",
             col.names = c("Month", "Date", "FSA", "Location ID", 
                           "Actual Capacity", "Funding Capacity", 
                           "Occupied Beds", "Unoccupied Beds", 
                           "Unavailable Beds", "Occupancy Rate"))
```
